Modeling site-specific amino-acid preferences deepens phylogenetic divergence estimates for viral proteins

### 2018-04-16 manuscript/manuscript.tex
removed period from title
added draft of empirical scatterplots

\title{Modeling site-specific amino-acid preferences deepens phylogenetic divergence estimates for viral proteins.}

@@

@@

However, while adding $\Gamma\omega$ rate variation increases the length of deep branches in a roughly uniform fashion across the tree, the branch lengthening from adding site-specific amino-acid preferences is not uniform across the tree (\ref{fig:empirical_trees}) \jdbcomment{I think we should make a Figure 5 that looks similar to the current Figure 6B, that would show GY94 versus each ExpCM, GY94+gamma versus each ExpCM+gamma, and each model versus it's gamma. These could be panels (one or two rows) and panel B.} 
Instead, the increase in branch length is most pronounced on branches leading to the HA sequence that was used in the deep mutational scanning experiment that informed the ExpCM.
For instance, the ExpCM informed by the H1 data most dramatically lengthens branches near the H1 clade of the tree, while the ExpCM informed by the H3 data has the largest effect on branches near the H3 clade.
The ExpCM informed by the average of the H1 and H3 data has a more uniform effect across the tree, but still most strongly extends branches leading to either the H1 or H3 clade.
Therefore, \ref{fig:empirical_trees} shows that ExpCMs estimate longer branches, but that the effect is shaped by the set of amino-acid preferences used to inform the model.

>>>>

\title{Modeling site-specific amino-acid preferences deepens phylogenetic divergence estimates for viral proteins}

@@

\begin{figure}
\centerline{\includegraphics[width=0.7\textwidth]{figures/empirical_scatter.pdf}}
\caption{\label{fig:empirical_scatter}
\textbf{Effect of site-specific amino-acid preferences and $\Gamma\omega$ rate variation on HA branch length estimation.} 
Points denote the branch lengths between all pairs of tips on the tree. 
Blue and orange denote branches that lead to the H1 and H3 deep mutational scanning reference sequences respectively. 
Each column has ExpCMs defined by a different deep mutational scanning amino-acid preference set. 
(A) The ExpCM defined by the H1, H3, or (H1+H3) average deep mutational scanning amino-acid preferences estimated longer branches from reference sequence used in the deep mutational scanning experiment. 
(B) The addition of $\Gamma\omega$ rate variation leads to more uniform extension of branch length than the addition of site-specific amino-acid preferences.
}
\end{figure}

@@

However, while adding $\Gamma\omega$ rate variation increases the length of deep branches in a roughly uniform fashion across the tree, the branch lengthening from adding site-specific amino-acid preferences is not uniform across the tree (\ref{fig:empirical_scatter}) 
Instead, the increase in branch length is most pronounced on branches leading to the HA sequence that was used in the deep mutational scanning experiment that informed the ExpCM.
For instance, the ExpCM informed by the H1 data most dramatically lengthens branches near the H1 clade of the tree, while the ExpCM informed by the H3 data has the largest effect on branches near the H3 clade.
The ExpCM informed by the average of the H1 and H3 data has a more uniform effect across the tree, but still most strongly extends branches leading to either the H1 or H3 clade.
Therefore, \ref{fig:empirical_trees} and \ref{fig:empirical_scatter} show that ExpCMs estimate longer branches, but that the effect is shaped by the set of amino-acid preferences used to inform the model.
